Coherent optical creation of a single molecule

### 2020-11-24 paper.tex
WIP: comments from Kang-Kuen

\begin{document}
\title{Coherent optical association of atoms into a single molecule}
% \title{Coherent optical creation of a single molecule}
\author{Yichao~Yu}
\email{yichaoyu@g.harvard.edu}
\gradstudent
\author{Kenneth~Wang}
\gradstudent
\author{Jonathan~D.~Hood}
\affiliation{Department of Chemistry, Purdue University, West Lafayette, Indianna, 47906, USA}
\author{Lewis~R.~B.~Picard}
\gradstudent
\author{Jessie~T.~Zhang}
\gradstudent
\author{William~B.~Cairncross}
\harvardccb
\harvardphysics
\cua
\author{Jeremy~M.~Hutson}
\affiliation{Joint Quantum Centre Durham-Newcastle, Department of Chemistry, Durham University, Durham, DH1 3LE, United Kingdom}
\author{Till Rosenband}
\harvardphysics
\author{Kang-Kuen~Ni}
\email{ni@chemistry.harvard.edu}
\harvardccb
\harvardphysics
\cua

@@

%% Preparation
Experimentally, we first prepare two atoms in a well-defined external and internal quantum state
using techniques developed previously~\cite{Liu2018, Liu2019, Wang2019}.
In brief, the experimental cycle begins by stochastically loading a single ${}^{23}\Na$ atom
and a single ${}^{133}\Cs$ atom into separate optical tweezers.
The atoms are initially imaged to distinguish between loading of two atoms,
one atom (Na or Cs), or no atom to be able to post-select from the experimental results
based on the initial loading condition.
Raman sideband cooling is then applied to prepare both atoms simultaneously
in the 3-dimensional motional ground state of their optical tweezers.
After cooling, the Na and Cs atoms are in the spin state~$\ket{\uparrow_{\Na}\uparrow_{\Cs}}\equiv \ket{f=2,m_f=2}_{\Na}\ket{f=4,m_f=4}_{\Cs}$,
which has a small scattering length.
The weak two-atom interaction allows the merging of the two tweezers to be done
with minimum pertubation so that they remain in the motional ground state.

@@

We choose the tweezer frequency to be far detuned (by $145 ~\mathrm{GHz}$)
from the $v' = 0$ line. Guided by the coupled-channel calculations,
we locate the Raman resonance for the atoms-to-molecule transition at
$770.57150(9) ~\mathrm{MHz}$ with about $3.75~\mathrm{mW}$ tweezer power focused down to a waist of $ 0.9~\mathrm{\mu m}$,
as shown in Fig.~\ref{f-raman}a.
% (We can maybe add information about the prediction here?)
%% Prediction
% This excited state used in the Raman transition was measured in our previous experiment using photoassociation to be at $288560 GHz$ from our atomic state. The ground molecular state has not been observed previously in experimentally. Based on our measurement of FB resonance, interaction shift and the binding energy of the 4422 bound state. Theory prediction was at $770.1 MHz$.
% The background level of $31~\mathrm{\%}$ corresponds to the probability of preparing the two atoms in the relative motional ground state.
% When the atoms are transferred into the molecule state by the Raman transition, there is a decrease in the two body survival since the resulting molecule is dark to our imaging sequence. % directly detected by our imaging step.
The molecular state is dark to the imaging step,
so successful transfer of the atoms to the molecular state is detected as loss of atoms.
We observe a Fourier-limited linewidth, which is evidence of a coherent transfer.
In order to verify the coherence of the transfer directly,
we fix the two-photon detuning on resonance and scan the pulse time.
Fig.~\ref{f-raman}b shows the observed coherent Rabi oscillation between the atomic and molecular states.
Fitting the data with a decaying Rabi oscillation suggests that
$69~\mathrm{\%}$ of atoms initially in the motional ground state are transferred into the molecular state after a $\pi$ pulse.
This transfer efficiency is limited mainly by the molecular lifetime,
which can be measured directly by preparing the molecule with a $\pi $ pulse
and then using a second $\pi$ pulse to dissociate the molecule back to atoms
after a variable wait time.
The result in Fig.~\ref{f-lifetime}a shows a molecular lifetime of $0.199(9)~\mathrm{ms}$
consistent with the decay of the Rabi oscillation.
We obtain the Raman Rabi frequency by fitting our measurements to a model
that includes a Raman Rabi frequency and
a finite lifetime for the molecular state, as shown in Fig.~\ref{f-raman}a and b.
We account for the small effect of atomic state loss by measuring the one-body and
two-body lifetime of the atoms directly without turning on the second frequency,
as shown in Fig.~\ref{f-lifetime}b.
The fit including all these effects gives a Raman Rabi frequency of $2\pi\times3.28(4)~\mathrm{kHz}$.

The efficiency of the transfer is lower than expected
because the ratio of the molecular scattering rate to the Rabi frequency
is 10 times larger than predicted. If this discrepancy arises from the $v'=0$ excited state,
it can be due either to a high ratio $\Omega_m / \Omega_a$ or to a large $\Gamma_e$.
Additionally, coupling to other excited states can add an offset to
both the Raman Rabi frequency and the scattering rate and thus affect their ratio.

>>>>

\begin{document}
% \title{Coherent optical association of atoms into a single molecule}
\title{Coherent optical creation of a single molecule}
\author{Yichao~Yu}
\email{yichaoyu@g.harvard.edu}
\gradstudent
\author{Kenneth~Wang}
\gradstudent
\author{Jonathan~D.~Hood}
\affiliation{Department of Chemistry, Purdue University, West Lafayette, Indianna, 47906, USA}
\author{Lewis~R.~B.~Picard}
\gradstudent
\author{Jessie~T.~Zhang}
\gradstudent
\author{William~B.~Cairncross}
\harvardccb
\harvardphysics
\cua
\author{Jeremy~M.~Hutson}
\affiliation{Joint Quantum Centre Durham-Newcastle, Department of Chemistry, Durham University, Durham, DH1 3LE, United Kingdom}
\author{Till Rosenband}
\harvardphysics
\author{Kang-Kuen~Ni}
\email{ni@chemistry.harvard.edu}
\harvardccb
\harvardphysics
\cua

@@

%% Preparation
Experimentally, we first prepare two atoms in a well-defined external and internal quantum state
using techniques developed previously~\cite{Liu2018, Liu2019, Wang2019}.
In brief, the experimental cycle begins by stochastically loading a single ${}^{23}\Na$ atom
and a single ${}^{133}\Cs$ atom into separate optical tweezers.
The atoms are initially imaged to distinguish between loading of two atoms,
one atom (Na or Cs), or no atom to be able to post-select from the experimental results
based on the initial loading condition.
After imaging, we turn on a $8.83(2)~\mathrm{G}$ magnetic field to define the quantization axis
for the state preparation and molecule formation steps.
Raman sideband cooling is then applied to prepare both atoms simultaneously
in the 3-dimensional motional ground state of their optical tweezers.
After cooling, the Na and Cs atoms are in the spin state~$\ket{\uparrow_{\Na}\uparrow_{\Cs}}\equiv \ket{f=2,m_f=2}_{\Na}\ket{f=4,m_f=4}_{\Cs}$,
which has a small scattering length.
The weak two-atom interaction allows the merging of the two tweezers to be done
with minimum pertubation so that they remain in the motional ground state.

@@

We choose the tweezer frequency to be $288560~\mathrm{GHz}$
which is far detuned (by $\approx\!145~\mathrm{GHz}$) from the $v' = 0$ line.
Guided by the coupled-channel calculations,
we locate the Raman resonance for the atoms-to-molecule transition at
$770.57150(9) ~\mathrm{MHz}$ with about $3.75~\mathrm{mW}$ tweezer power
focused down to a waist of $ 0.9~\mathrm{\mu m}$, as shown in Fig.~\ref{f-raman}a.
% (We can maybe add information about the prediction here?)
%% Prediction
% This excited state used in the Raman transition was measured in our previous experiment using photoassociation to be at $288560 GHz$ from our atomic state. The ground molecular state has not been observed previously in experimentally. Based on our measurement of FB resonance, interaction shift and the binding energy of the 4422 bound state. Theory prediction was at $770.1 MHz$.
% The background level of $31~\mathrm{\%}$ corresponds to the probability of preparing the two atoms in the relative motional ground state.
% When the atoms are transferred into the molecule state by the Raman transition, there is a decrease in the two body survival since the resulting molecule is dark to our imaging sequence. % directly detected by our imaging step.
The molecular state is dark to the imaging step,
so successful transfer of the atoms to the molecular state is detected as loss of atoms.
We observe a Fourier-limited linewidth, which is evidence of a coherent transfer.
In order to verify the coherence of the transfer directly,
we fix the two-photon detuning on resonance and scan the pulse time.
Fig.~\ref{f-raman}b shows the observed coherent Rabi oscillation between the atomic and molecular states.
Fitting the data with a decaying Rabi oscillation suggests that
$69~\mathrm{\%}$ of atoms initially in the motional ground state are transferred into the molecular state after a $\pi$ pulse.
This transfer efficiency is limited mainly by the molecular lifetime,
which can be measured directly by preparing the molecule with a $\pi $ pulse
and then using a second $\pi$ pulse to dissociate the molecule back to atoms
after a variable wait time.
The result in Fig.~\ref{f-lifetime}a shows a molecular lifetime of $0.199(9)~\mathrm{ms}$
consistent with the decay of the Rabi oscillation.
We obtain the Raman Rabi frequency by fitting our measurements to a model
that includes a Raman Rabi frequency and
a finite lifetime for the molecular state, as shown in Fig.~\ref{f-raman}a and b.
We account for the small effect of atomic state loss by measuring the one-body and
two-body lifetime of the atoms directly without turning on the second frequency,
as shown in Fig.~\ref{f-lifetime}b.
The fit including all these effects gives a Raman Rabi frequency
of $2\pi\times3.28(4)~\mathrm{kHz}$.

The efficiency of the transfer is lower than expected
because we found the ratio of the molecular scattering rate to the Rabi frequency
is 10 times larger than predicted. If this discrepancy arises from the $v'=0$ excited state,
it can be due either to a high ratio $\Omega_m / \Omega_a$ or to a large $\Gamma_e$.
Additionally, coupling to other excited states can add an offset to
both the Raman Rabi frequency and the scattering rate and thus affect their ratio.
